Graphical Tests for Power Comparison of Competing Designs

### 2012-03-31 experiment-two.tex
small changes
Merge branch 'master' of github.com:heike/comparing-designs
Conflicts:
	setup.tex

% !TEX root = comparison.tex

\subsection{Comparing Distributions}
\begin{figure*}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=\linewidth]{power-exp2} 
   \caption{Overview of power predictions for the four different designs. The fully saturated thick lines show average predicted power for each of the designs facetted by size of the red group (top to bottom) and relative size of the blue group to the red group (left to right). Thin lines represent variability due to subject-specific abilities. }
   \label{fig:power2}
\end{figure*}

@@

The two highly summarized designs benefit the most from an increase in sample size: both density charts and boxplots profit significantly more than histograms and dotplots. This might, again, be a hint towards a trade-off between visual complexity and amount of information shown as basis for decisions.

Table \ref{tbl:speed2} gives an overview of results from a linear model for (log) evaluation time of lineups. Subject-specific speed (due to spatial ability) are accounted for  including a random intercept. Structurally, the model is the same as before, i.e. we regard effect size $d$, size $n_1$ of the first group, and the ratio $n_2/n_1$ between group sizes both as main effects and in two-way interactions with the four competing designs. To make estimate sizes comparable, we have again multiplied all estimates and standard errors involving the group size $n_1$ by 100. This means that we need to interpret those  estimates as average change if the group size is increased by 100 elements. 
The results of this model are similar to the previous: evaluation speeds needed for the four different designs is very close, but  histograms take on average a significantly longer time ($e^{0.23} = 1.26$ seconds) to evaluate than lineups of boxplots. 
Where an increase of sample size led to an increase of accuracy before, we now see that it increases time for evaluation of boxplot slightly, but not significantly so. What is interesting is that evaluation time decreases significantly with an increase in sample size, particularly for dotplots, but also for density charts. This is counter-intuitive to our previous argument on dotplots being affected by visual complexity the most. It seems, that observers of dotplot lineups come to a conclusion faster as sample size increases, albeit more often to  the wrong conclusion. Rather than visual complexity, this might be indicative that over-plotting is problematic in dotplots. While we tried to accommodate for over-plotting by making use of alpha-blending and jittering, this might not have been sufficient for the large sample size.

@@

>>>>

% !TEX root = comparison.tex
\subsection{Comparing Distribution Centers}

@@

The two highly summarized designs benefit the most from an increase in sample size: both density charts and boxplots profit significantly more than histograms and dotplots. This might, again, be a hint towards a trade-off between visual complexity and amount of information shown as the basis for making decisions.

Table \ref{tbl:speed2} gives an overview of results from a linear model for (log) evaluation time of lineups. Subject-specific speed (due to spatial ability) are accounted for  including a random intercept. Structurally, the model is the same as before, i.e. we regard effect size $d$, size $n_1$ of the first group, and the ratio $n_2/n_1$ between group sizes both as main effects and in two-way interactions with the four competing designs. To make estimate sizes comparable, we have again multiplied all estimates and standard errors involving the group size $n_1$ by 100. This means that we need to interpret those  estimates as average change if the group size is increased by 100 elements.

@@

The results of this model are similar to the previous: evaluation speeds needed for the four different designs is very close, but  histograms take on average a significantly longer time ($e^{0.23} = 1.26$ seconds) to evaluate than lineups of boxplots. 
Where an increase of sample size led to an increase of accuracy before, we now see that it increases time for evaluation of boxplot slightly, but not significantly so. What is interesting is that evaluation time decreases significantly with an increase in sample size, particularly for dotplots, but also for density charts. This is counter-intuitive to our previous argument on dotplots being affected by visual complexity the most. It seems, that observers of dotplot lineups come to a conclusion faster as sample size increases, albeit more often to  the wrong conclusion. Rather than visual complexity, this might be indicative that over-plotting is problematic in dotplots. While we tried to accommodate for over-plotting by making use of alpha-blending and jittering, this might not have been sufficient for the large sample size.
